---
title: "Group3 sales forecasting"
author: "Group3"
date: "2024-06-13"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r}
if (!requireNamespace("dplyr", quietly = TRUE)) {
  install.packages("dplyr")
}
if (!requireNamespace("ggplot2", quietly = TRUE)) {
  install.packages("ggplot2")
}
if (!requireNamespace("readxl", quietly = TRUE)) {
  install.packages("readxl")
}

library(dplyr)
library(ggplot2)
library(readxl)

# Load the dataset
salesforecasting <- read_excel("C:\\Users\\srira\\Downloads\\salesforecasting.xlsx")
```

About Dataset
This dataset offers a valuable resource for businesses operating in the retail furniture sector. By analyzing historical sales data from the superstore dataset, users can gain insights into future sales patterns and trends. This information can be utilized to optimize inventory management strategies, anticipate customer demand, and enhance overall operational efficiency. Whether for retail managers, analysts, or data scientists, this dataset provides a foundation for informed decision-making, helping businesses maintain stability and drive sustained growth in the dynamic retail environment.

```{r Print the structure of your dataset}
str(salesforecasting)
```


```{r List the variables in your dataset}
names(salesforecasting)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
```{r Print the top 15 rows of your dataset}
head(salesforecasting, n=15)
```
```{r Write a user defined function using any of the variables from the data set.}
calculate_total_sales_by_subcategory <- function(df, subcategory) {
  total_sales <- df %>%
    filter(salesforecasting$`Sub-Category` == subcategory) %>%
    summarize(TotalSales = sum(Sales, na.rm = TRUE))
  return(total_sales)
}

# Call the function with the dataset and the sub-category "Chairs"
total_sales_chairs <- calculate_total_sales_by_subcategory(salesforecasting, "Chairs")
print(total_sales_chairs)
```
```{r Use data manipulation techniques and filter rows based on any logical criteria that exist in your dataset.}
# Define a simple user-defined function to filter the data by state and calculate total profit
total_profit_by_state <- function(df, state) {
  # Filter the data for the given state
  filtered_data <- df %>% filter(salesforecasting$State == state)
  
  # Calculate the total profit for the filtered data
  total_profit <- sum(filtered_data$Profit, na.rm = TRUE)
  
  return(total_profit)
}

# Call the function with the dataset and the state "California"
total_profit_california <- total_profit_by_state(salesforecasting, "California")
print(total_profit_california)

```
```{r Identify the dependent & independent variables and use reshaping techniques and create a new data frame by joining those variables from your dataset.}
# Select relevant columns for the new data frame
selected_columns <- salesforecasting %>% select(`Order ID`,`Customer ID`,Segment,Category,`Sub-Category`,Sales,Quantity,Discount,Profit)

# Summarize the data by Segment and Category
summary_df <- selected_columns %>%
group_by(Segment, Category) %>%
  summarize(Total_Profit = sum(Profit, na.rm = TRUE),
            Average_Sales = mean(Sales, na.rm = TRUE),
            Total_Quantity = sum(Quantity, na.rm = TRUE))

# Print the new data frame
print(summary_df)
```


```{r Remove missing values in your dataset.}
# Remove rows with missing values
if(any(is.na(data))) {
  # Remove rows with missing values
  data_clean <- data %>% na.omit()
  
  # Display the cleaned data
  print("Data after removing rows with missing values:")
  print(data_clean)
} else {
  print("No missing values")
}

```

```{r Identify and remove duplicated data in your dataset}
# Identify duplicated rows based on selected columns
duplicated_rows <- salesforecasting %>%
  group_by(`Order ID`, `Order Date`, `Ship Date`, `Customer ID`) %>%
  filter(!duplicated(`Order ID`))

# Print cleaned dataset
print(duplicated_rows)

```
```{r Reorder multiple rows in descending order}
# Reorder rows based on Sales in descending order
data_ordered <- salesforecasting %>%
  arrange(desc(Sales))

# Print the reordered data
print(data_ordered)

```

```{r Rename some of the column names in your dataset}
# Rename specific columns
data_renamed <- salesforecasting %>%
  rename(
    Row_ID = `Row ID`,
    Order_ID = `Order ID`,
    Order_Date = `Order Date`,
    Ship_Date = `Ship Date`,
  )

# Print the data with renamed columns
print(data_renamed)
```


```{r Add new variables in your data frame by using a mathematical function)}

# Add a new variable by multiplying an existing column by 2
data_modified <- salesforecasting %>%
  mutate(SalesTwice = Sales * 2)

# Print the modified data frame with the new variable
print(data_modified)
```

```{r Create a training set using random number generator engine}
set.seed(123)

# Number of rows in the training set
train_size <- 0.7 * nrow(salesforecasting)  # Adjust 0.7 to your desired proportion

# Generate indices for training set
train_indices <- sample(seq_len(nrow(salesforecasting)), size = train_size, replace = FALSE)

# Create training set
train_set <- salesforecasting[train_indices, ]

# Print the first few rows of the training set
print(head(train_set))
```

```{r # Print the summary statistics of your dataset}
summary(salesforecasting)

```

```{r Use any of the numerical variables from the dataset and perform the following statistical functions • Mean • Median • Mode • Range}
mean_sales <- mean(salesforecasting$Sales, na.rm = TRUE)
print(mean_sales)
median_sales <- median(salesforecasting$Sales, na.rm = TRUE)
print(median_sales)
mode_sales <- as.numeric(names(sort(table(salesforecasting$Sales), decreasing = TRUE)[1]))
print(mode_sales)
range_sales <- range(salesforecasting$Sales, na.rm = TRUE)
print(range_sales)

```

```{r Plot a scatter plot for any 2 variables in your dataset}
ggplot(salesforecasting, aes(x = Quantity, y = Sales)) + geom_point() + labs(title = "Scatter Plot of Quantity vs Sales")
```

```{r Plot a bar plot for any 2 variables in your dataset}
#ggplot(salesforecasting, aes(x = Category, y = Sales)) + geom_bar(stat = "identity") + labs(title = "Bar Plot of Category vs Sales")
ggplot(salesforecasting, aes(x = `Sub-Category`, y = Sales)) +
  geom_bar(stat = "identity") +
  labs(title = "Total Sales by Sub-Category", x = "Sub-Category", y = "Total Sales") +
  theme_minimal()
```

```{r Find the correlation between any 2 variables by applying Pearson correlation}
correlation_sales_profit <- cor(salesforecasting$Sales, salesforecasting$Profit, method = "pearson")

# Print the correlation
print(correlation_sales_profit)
```

